\documentclass[11pt]{article}
\pagestyle{empty}
\usepackage{color}
\usepackage{listings}
\usepackage{minted}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[options]{algorithm2e}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\cfoot[R]{\thepage~of~\pageref{LastPage}}
\addtolength{\oddsidemargin}{-.875in}
\addtolength{\evensidemargin}{-.875in}
\addtolength{\textwidth}{1.75in}
\addtolength{\topmargin}{-.875in}
\addtolength{\textheight}{1.75in}

\begin{document}
\begin{center} {\Large\bf FA, Homework 8} \\ Quentin McGaw (qm301) \\ 04/06/17
\end{center}

\begin{quote}
The search for truth is more precious than its possession
\\ --  Albert Einstein
\end{quote}

\begin{enumerate}
\item \textbf{\textcolor{blue}{Set $W=\lfloor{\sqrt{N}}\rfloor$.  
We are {\em given} $PRICE[I]$, $1\leq I\leq W$, the price of a rod of length $I$. 
Give a program that will output the optimal revenue for a rod of length $N^2$ 
and give the time, in $\Theta$-land,  of the algorithm.
Use an auxiliary array $R[J]$, $0\leq J\leq N^2$, where $R[J]$ will give the
optimal revenue for a rod of length $J$.  You may {\bf not} use
the term {\tt MAX} nor {\tt MIN} in your program  Explain, in clear words, how your program is working.
(You can and should use {\tt MAX} in your explanations.)}}
    \\\\ We have $R[J] = max[PRICE[I] + R[J-I]]$ where $1 \leq I \leq min[I, W]$. \\\\
    \begin{algorithm}[H]
        R[0] = 0 \\
        R[1] = PRICE[1] \\
        \For{J = $2$ to $N^2$}{
            S = J \\
            \If{W $\leq$ S}{
                S = W \\
            }
            \\ (* So S = min(J,W) *) \\
            TEMP = 0 \\
            \For{I = 1 to S}{
                \If{TEMP $\leq$ PRICE[I] + R[J-I]}{
                    TEMP $\leftarrow$ PRICE[I] + R[J - I] \\          
                }
            }
            \\ (* Here TEMP is the maximal value *) \\
            R[j] $\leftarrow$ TEMP \\
        }
        \Return{R[$N^2$]}
        \caption{Algorithm to find the optimal revenue for a rod of length I}
    \end{algorithm}
    \\ The outer loop executes $1 \leq J \leq N^2$ times.
    \\ The time of the inner loop is min(J,W) and there are thus two ranges:
    \begin{itemize}
        \item $1 \leq J < W$ : J steps and it adds up to $1 + 2 + ... + (W - 1) \approx \frac{W^2}{2} = \Theta(N)$.
        \item $W \leq J \leq N^2$ : W steps and it adds up to W($N^2$ - W) = $\Theta(N^{2.5})$.
    \end{itemize}
    The total time is therefore $\Theta(N + N^{2.5}) = \Theta(N^{2.5})$.
    
\item  \textbf{\textcolor{blue}{Suppose, in the Activity Selector problem, we
instead select the last activity to start that is compatible with all previously
selected activities. Describe how this approach works, write a program for it 
(pseudo-code allowed) and prove that it yields an optimal algorithm.}}
    \\ At each point, the last activity to start is selected. There is a recursion down to the single optimal subproblem to find the optimal solution for all remaining activities compatible (proof by induction). \\
    \begin{algorithm}[H]
    RECURSIVE-ACTIVITY-SELECTOR(s, f, i): \\
    $m \leftarrow i - 1$ \\
    \While{m $>$ 0 and $s_i < f_m$}{
        m $\leftarrow$ m - 1   
    }
    \eIf{m $>$ 0}{
        \Return{$\{a_m\}\ \cup $ RECURSIVE-ACTIVITY-SELECTOR(s, f, m)}
    }{
        \Return{$\emptyset$}
    }
    \caption{Recursive activity selector}
    \end{algorithm}
    
    \\ We start with i = n (number of tasks). Consider a nonempty subproblem $S_{ij}$ and let $a_m$ be the activity in $S_{ij}$ with the latest start time $s_m = \max(s_k : a_k \in S_{ij})$.
    \begin{itemize}
        \item $a_m$ is used in a maximum-size subset of mutually compatible activities $S_{ij}$.
        \\ \textbf{Proof: } Suppose $A_{ij}$ is a maximum-size subset of mutually compatible activies of $S_{ij}$.
        Suppose activities in $A_{ij}$ are ordered in monotonically increasing order of starting time.
        Let $a_k$ be the last activity of $A_{ij}$. If $a_k = a_m$, we are done since $a_m$ is used in constructing the schedule. Otherwise, we construct $A^{'}_{ij} = A_{ij} - \{a_k\}\ \in \ \{a_m\}$. The activities are disjoint since $a_k$ is the last activity to start and $s_m \geq s_k$. The number of activities in the subset are the same, so it is a maximum-size subset of activities including $a_m$.
        \item The subproblem $S_{mj}$ is empty, so choosing $a_m$ leaves the subproblem $S_im$ as the only one that may be nonempty.
        \\ \textbf{Proof: } Suppose $S_{mj}$ is non-empty, so there is an activity $a_k$ such that $f_m \leq s_k < f_k \leq s_j < f_j$. As a consequence, $a_k \in S_{ij}$ which has a later start time than $a_m$, contradicting our initial assumption.
    \end{itemize}
    


    \\ Let $S$ be the set of activities and $A$ be the maximum-size subset of mutually compatible activities of $S$. We first arrange the activities in order of decreasing start time. Let $a_k = [s_k, f_k)$ be the last activity in $A$. Let $a_m$ be the activity with the earliest finish time. At every step, we discard any overlapping activities with already chosen activities. The latest starting activities remains chosen.
    \\\\ If $a_k = a_m$ then we found an optimal solution
    \\ Else, we build the new set starting with $a_k$ and go up until $a_m \leq a_k$. This set has the same number of activities and is mutually compatible with the original set. The optimal solution for $S$ maps directly to the optimal solution in the new set.    

\item \textbf{\textcolor{blue}{Students (professors too!) often come up with very clever ideas for
optimization programs.  The problem (often!) is that they (sometimes, but that
is enough) give the wrong answer.  Here are three approaches and {\em your} problem,
in each case, is to give an example where it yields the wrong answer.}}
    \begin{enumerate}
    \item \textbf{\textcolor{blue}{Pick the activity of the shortest duration 
    from amongst those which do not overlap previously selected activities.}}
        \\ With the set of activities $A = \{a_1, a_2, a_3\}$ where their start times are $s = \{2, 0, 3\}$ and finish times are $f = \{4, 3, 6\}$. Then their respective durations are $d = \{2, 3, 3\}$.
        \\ The solution obtained would be $\{a_1\}$ whilst the optimal solution is actually $\{a_2, a_3\}$.
    \item \textbf{\textcolor{blue}{(*) Pick the activity which overlaps the fewest other remaining activities
    from amongst those which do not overlap previously selected activities.}}
        \\ With the set of activities $A = \{a_1, a_2, a_3, ..., a_9\}$ where their start times are $s = \{0, 1, 1, 2, 3, 4, 5, 5, 6\}$ and finish times are $f = \{2, 3, 3, 4, 5, 6, 7, 7, 8\}$. Then their respective number of overlaps are $L = \{2, 3, 3, 3, 2, 3, 3, 3, 2\}$
        \\ The solution obtained would be $\{a_1, a_5, a_9\}$ whilst the optimal solution is actually $\{a_1, a_9, a_4, a_6\}$.
    \item \textbf{\textcolor{blue}{Pick the activity with the earliest start time
    from amongst those which do not overlap previously selected activities.}}
        \\ With the set of activities $A = \{a_1, a_2, a_3\}$ where their start times are $s = \{0, 1, 3\}$ and finish times are $f = \{4, 2, 4\}$.
        \\ The solution obtained would be $\{a_1\}$ whilst the optimal solution is actually $\{a_2, a_3\}$.
    \end{enumerate}
\end{enumerate}


\begin{quote}
Dear Sir,
\\  I beg to introduce myself to you as a clerk in the Accounts Department
of the Port Trust Office at Madras on a salary of only \pounds 20 per
annum.  I am now about 25 years of age.  I have no University education but
I have undergone the ordinary school course.  After leaving school I have
been employing the spare time at my disposal to work at Mathematics\ldots
I am striking out a new path for myself.  I have made a special investigation
of divergent series in general and the results I get are termed
by the local mathematicians as ``startling"  \\ -- Ramanujan

\end{quote}

\begin{quote}
The truth is that the theory of primes is full of pitfalls, to surmount which requires the 
fullest of trainings in modern rigorous methods. This you are naturally without. I hope 
you will not be discouraged by my criticisms. I think your argument a very remarkable 
and ingenious one. To have proved what you claimed to have proved would have been
 about the most remarkable mathematical feat in the whole history of mathematics.
\\ G.H. Hardy, letter to Ramanujan, 1913 [after finding error in Ramanujan argument]
\end{quote}

\end{document}

